---
title: "Draft Quarto document"
author: "Fie Langmann"
format: html
---

## Header 1

```{r setup}
#| include: false
targets::tar_config_set(store = here::here("_targets"))
library(tidyverse)
library(targets)
library(tidymodels)
source(here::here("R/functions.R"))
lipidomics <- targets::tar_read(lipidomics)
```

## Results

```{r}
targets::tar_read(df_stats_by_metabolite) |>
  mutate(MeanSD = glue::glue("{value_mean} ({value_sd})")) |>
  select(Metabolite = metabolite, `Mean SD` = MeanSD) |>
  knitr::kable(caption = "Descriptive statistics of the metabolites.")
```

```{r}
targets::tar_read(fig_metabolite_distribution)
```

## Building the model

Working with logistic regression for this research question:

*What is the estimated relationship of each metabolite with T1D compared
to the controls, adjusting for the influence of age and gender?*

```{r}
log_reg_specs <- logistic_reg() |>
  set_engine("glm") # using general linear models
log_reg_specs
```

We need to do some data transformation. Data is on a long format and
there are multiple Cholesterol values, while all other metabolites only
have one. The last issue is that each metabolite has quite large
differences in the values and ranges of data. In our research question
we want to know how each metabolite influences T1D. In order to best
interpret the results and compare across metabolites, we should ideally
have all the metabolites with a similar range and distribution of values

we can tidy up the metabolite names using mutate() with
snakecase::to_snake_case(). Secondly, to make our dataset wider, we use
pivot_wider(). Since we also want an easy way of identifying columns
that are metabolites, we will add a "metabolite\_" prefix using the
argument names_prefix. To fix the multiple cholesterol issue, we will
merge the values by calculating a mean before pivoting. We do this by
setting the values_fn with mean.

```{r}
lipidomics |>
  dplyr::count(code, metabolite) |>
  dplyr::filter(n > 1)
```

Since we’re using a function-oriented workflow, let’s convert both the
“metabolite to snakecase” and “pivot to wider” code into their own
functions in the R/functions.R file

```{r}
lipidomics_wide <- lipidomics |>
  column_values_to_snake_case(metabolite) |>
  metabolites_to_wider()

lipidomics_wide
```

Now we will fix the differences in scales for the metabolites

```{r}
recipe(lipidomics_wide) |>
  update_role(metabolite_cholesterol, age, gender, new_role = "predictor") |>
  update_role(class, new_role = "outcome") |>
step_normalize(starts_with("metabolite_"))
```

```{r}
recipe_specs <- lipidomics_wide |>
  create_recipe_spec(metabolite_cholesterol)
recipe_specs
```
